---
title: "ML2 Team 1 Presentation 2"
output: html_document
---

## Linear Model Selection and Regularization {.tabset .tabset-fade} 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Data

</br></br>

Test Will be used for Validation
Train will be Split in .8/.2 for model testing. 

```{r, collapse=T, comment=NA, warning=F}
Trans <- read.csv("D:/Downloads/train_transaction.csv")
head(Trans, 20)
colnames(Trans)
```
A lot of columns-  No good desc on what many of them do

```{r, collapse=T, comment=NA, warning=F}
Trans<-Trans[,1:55] # cut out "V" fields
Iden <- read.csv("D:/Downloads/train_identity.csv")
head(Iden,20)
```

**Adding Info for device.  Get most comon Device type**

```{r, collapse=T, comment=NA, warning=F}
library(dplyr)
library(tidyr)
```
From Top team -> card1,addr1, and D1
```{r, collapse=T, comment=NA, warning=F}
User<-Trans %>% select(TransactionID,card1, addr1, D1) 
Device<- Iden %>% select(TransactionID,DeviceType,DeviceInfo)
```
we now have linking fields on transaction ID to join Identity to the User Table
```{r, collapse=T, comment=NA, warning=F}
SummaryTable_USER_DEVICE <-left_join(User, Device, by = "TransactionID")
```
Replacing null values with "Unknown"  
```{r, collapse=T, comment=NA, warning=F}
SummaryTable_USER_DEVICE<- SummaryTable_USER_DEVICE %>% mutate(DeviceInfo = replace_na(DeviceInfo, "UnKNOWN")) %>%
  mutate(DeviceInfo= replace(DeviceInfo,DeviceInfo=="", "UnKNOWN")) %>% mutate(DeviceType = replace_na(DeviceType, "UnKNOWN")) %>%
  mutate(DeviceType= replace(DeviceType,DeviceType=="", "UnKNOWN"))
```
Creating a surrogate key for user id 
```{r, collapse=T, comment=NA, warning=F}
OneKey<- filter(SummaryTable_USER_DEVICE,DeviceInfo!="UnKNOWN") %>% unite(User,'card1','addr1','D1',sep='|')%>% 
  unite(Device,'DeviceType','DeviceInfo',sep='|') 
```
Grouping by the surrogate key , couting the occurrence of device and taking the top one per user id 
```{r, collapse=T, comment=NA, warning=F}

MostCommon<- OneKey %>% group_by(User)%>% count(Device) %>% slice(which.max(1))

```
Taking only the linking data and # user 
```{r, collapse=T, comment=NA, warning=F}
MostCommon<- MostCommon %>% select('User','Device')
```
Joining data back to most common and one key which has trasnaction ID
```{r, collapse=T, comment=NA, warning=F}
Joined<-  left_join(OneKey, MostCommon, by = "User")
```
Seeing if the device in the transaction matches the most common for the user
```{r, collapse=T, comment=NA, warning=F}
TranMatched<- Joined %>% mutate(Matched= (Device.x==Device.y))
```
Taking only the transaction ID and the resulting boolean
```{r, collapse=T, comment=NA, warning=F}
TranMatched<- TranMatched%>% select('TransactionID','Matched')
```

**Getting Back to Transactions**
```{r, collapse=T, comment=NA, warning=F}
Trans$card4<-as.factor(Trans$card4) #Master Visa Etc. 

Trans$isFraud<-as.factor(Trans$isFraud) #Binary Class

Trans$ProductCD <-as.factor(Trans$ProductCD ) #Product type
```
Creating Boolean for email domain matching or True if second domain is blank (bill to and send to are the same)
```{r, collapse=T, comment=NA, warning=F}
WorkingTrans<- Trans  %>%
  mutate(EmailsMatch=((P_emaildomain)==(R_emaildomain)| R_emaildomain==''))
```
Convert to factor
```{r, collapse=T, comment=NA, warning=F}
WorkingTrans$EmailsMatch<- as.factor(WorkingTrans$EmailsMatch)
```
Select fields
```{r, collapse=T, comment=NA, warning=F}

FirstDataSet<- WorkingTrans %>% select('TransactionID','isFraud','TransactionAmt','ProductCD','card1','card2','card3','card4','card5','card6'
                                       ,'addr1','addr2','EmailsMatch')
```
Join our matching information on device 

```{r, collapse=T, comment=NA, warning=F}
JoinedSet<- left_join(FirstDataSet, TranMatched, by = "TransactionID")
```
Matched as factor 
```{r, collapse=T, comment=NA, warning=F}
JoinedSet$Matched<- as.factor(JoinedSet$Matched)
```
Final output 
```{r, collapse=T, comment=NA, warning=F}
Output_SET_1<- JoinedSet %>% select('isFraud','TransactionAmt','ProductCD','card1','card2','card3','card4','card5','card6'
                                    ,'addr1','addr2','EmailsMatch','Matched')
```

### Ridge

</br></br>

```{r, collapse=T, comment=NA, warning=F}
library(glmnet)
```
Make sure is fraud is a factor for binary classification 

```{r, collapse=T, comment=NA, warning=F}
Output_SET_1$isFraud<-as.factor(Output_SET_1$isFraud)
```
Fixing sparse factors where small occurence types exist
```{r, collapse=T, comment=NA, warning=F}
Output_SET_1<-Output_SET_1%>% mutate(Matched = replace_na(Matched, FALSE)) %>% mutate(card6 = replace(card6,card6=='charge card','credit')) %>%
  mutate(card6 = replace(card6,card6=='debit or credit','debit')) %>%  mutate(card6 = replace(card6,card6=='','credit'))

summary(Output_SET_1)
```
Re casting it to factor
```{r, collapse=T, comment=NA, warning=F}
Output_SET_1$card6<-as.factor(as.character(Output_SET_1$card6)) #Credit/Debit
```
Perform mean replacement for other numeric fields if na
```{r, collapse=T, comment=NA, warning=F}
for(i in 1:ncol(Output_SET_1)){
  Output_SET_1[is.na(Output_SET_1[,i]), i] <- mean(Output_SET_1[,i], na.rm = TRUE)
}


summary(Output_SET_1)
```
Split data
```{r, collapse=T, comment=NA, warning=F}
set.seed(2021)
LenData<-nrow(Output_SET_1)
Test.Index<-sample(1:LenData*.8)
```
Create Train and Test Sets
```{r, collapse=T, comment=NA, warning=F}
Train<-Output_SET_1[Test.Index,]
Test<- Output_SET_1[-Test.Index,]
```

Create x Matrix
```{r, collapse=T, comment=NA, warning=F}
x_train <- model.matrix( ~ .-1, Train[,2:ncol(Train)])
summary(Output_SET_1)
```
Create label vector
```{r, collapse=T, comment=NA, warning=F}
Y.Train<-Train[,1]
```
Create test matrix and labels
```{r, collapse=T, comment=NA, warning=F}
x_test <- model.matrix( ~ .-1, Test[,2:ncol(Test)])
Y.Test<-Test[,1]
```
Create grid of lambdas

```{r, collapse=T, comment=NA, warning=F}
lambdas <- 10^seq(3, -2, by = -1)
```
Fit GLM
```{r, collapse=T, comment=NA, warning=F}
GLM_FIT <- cv.glmnet(x_train, Y.Train, alpha = 0, lambda = lambdas,family='binomial')
summary(GLM_FIT)

```
Get best Lambda 

```{r, collapse=T, comment=NA, warning=F}
best_lambda <- GLM_FIT$lambda[which.min(GLM_FIT$cvm)]
best_lambda
```
Create best model

```{r, collapse=T, comment=NA, warning=F}
Best_Ridge<- glmnet(x_train, Y.Train, alpha = 0, lambda = best_lambda,family='binomial')
```
Predict class (.5 threshold)
```{r, collapse=T, comment=NA, warning=F}
preds<-predict(Best_Ridge,newx = x_test,type = 'class')
table(preds,Y.Test)
```
Get raw prediction values
```{r, collapse=T, comment=NA, warning=F}
predsVal<-predict(Best_Ridge,newx = x_test)
```
Plot AUC 
```{r, collapse=T, comment=NA, warning=F}
library(ROCR)
Ridge_AUCThing <- prediction(predsVal,Y.Test, label.ordering = c(0,1))
Ridge_Chart <- performance(Ridge_AUCThing,"tpr", "fpr")
plot(Ridge_Chart)
```
Get AUC value
```{r, collapse=T, comment=NA, warning=F}
auc_ROCR <- performance(Ridge_AUCThing, measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
auc_ROCR # [1] 0.726611
```
### Lasso

</br></br>
Fit Lasso model
```{r, collapse=T, comment=NA, warning=F}
GLM_FIT_LASSO <- cv.glmnet(x_train, Y.Train, alpha = 1, lambda = lambdas,family='binomial',trainControl='cv')
summary(GLM_FIT_LASSO)
```
Best lambda
```{r, collapse=T, comment=NA, warning=F}
best_lambda <- GLM_FIT_LASSO$lambda[which.min(GLM_FIT_LASSO$cvm)]
best_lambda
```
Best model
```{r, collapse=T, comment=NA, warning=F}
Best_Lasso<- glmnet(x_train, Y.Train, alpha = 1, lambda = best_lambda,family='binomial')
```
Basic predictions
```{r, collapse=T, comment=NA, warning=F}
preds<-predict(Best_Lasso,newx = x_test,type = 'class')
table(preds,Y.Test)
```
Raw Predcitions 
```{r, collapse=T, comment=NA, warning=F}
predsVal_Lasso<-predict(Best_Lasso,newx = x_test)
```
Plot AUC
```{r, collapse=T, comment=NA, warning=F}
library(ROCR)
Lasso_AUCThing <- prediction(predsVal_Lasso,Y.Test, label.ordering = c(0,1))
Lasso_Chart <- performance(Lasso_AUCThing,"tpr", "fpr")
plot(Lasso_Chart)
```
Get AUC value 
```{r, collapse=T, comment=NA, warning=F}
auc_ROCR <- performance(Lasso_AUCThing, measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
auc_ROCR # [1] 0.7265185 # on new run 0.7165336
```

### XGBOOST

</br></br>
```{r, collapse=T, comment=NA, warning=F}
library(xgboost)
```
Create numeric labels for XGboost
```{r, collapse=T, comment=NA, warning=F}
labels = as.numeric(Train[,1])-1

xgboost_model <- xgboost(data = x_train, label = labels,max_depth = 18, 
                         objective = "binary:logistic", nrounds = 100, 
                         verbose = TRUE, prediction = TRUE)

summary(xgboost_model)
xgboost_model
```
Predict outcomes
```{r, collapse=T, comment=NA, warning=F}
boost_preds<-predict(xgboost_model,newdata = x_test,type='class')
```
Check .5 thres confusion matrix
```{r, collapse=T, comment=NA, warning=F}
table(boost_preds>=.5,Test$isFraud)
```
Plot AUC
```{r, collapse=T, comment=NA, warning=F}
library(ROCR)
XGBOOST_AUCThing <- prediction(boost_preds,Test$isFraud, label.ordering = c(0,1))
XGBOOST_Chart <- performance(XGBOOST_AUCThing,"tpr", "fpr")
plot(XGBOOST_Chart)

```
Get AUC value
```{r, collapse=T, comment=NA, warning=F}
auc_ROCR <- performance(XGBOOST_AUCThing, measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
auc_ROCR # [1] 0.7882256 second run -0.7851282
```

### GBM
</br></br>

```{r, collapse=T, comment=NA, warning=F}
library(caret)
```

Fit GBM
```{r, collapse=T, comment=NA, warning=F}
GBM <- train(isFraud~.,data = Train,method = "gbm",
             preProcess = c("scale", "center"),trControl = trainControl(method = "cv", 
                                                                        number = 3, verboseIter = TRUE),verbose = 0)

GBM
```
Find GBM Classes 

```{r, collapse=T, comment=NA, warning=F}
preds<-predict(GBM,newdata = Test)
table(preds,Test$isFraud)
```
Get GBM Probabilities
```{r, collapse=T, comment=NA, warning=F}
predsVal<-predict(GBM,Test,type='prob')
```
Take only positive value class values
```{r, collapse=T, comment=NA, warning=F}
Only<-predsVal[,2]
```
Plot AUC 
```{r, collapse=T, comment=NA, warning=F}

library(ROCR)
GBM_AUCThing <- prediction(Only,Test$isFraud, label.ordering = c(0,1))
GBM_Chart <- performance(GBM_AUCThing,"tpr", "fpr")
plot(GBM_Chart)
Gbm_Final<-GBM$finalModel
summary(Gbm_Final)
```
Get AUC value
```{r, collapse=T, comment=NA, warning=F}
auc_ROCR <- performance(GBM_AUCThing, measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
auc_ROCR # [1] 0.7619435

```

### Conclusion
</br></br>
Plots for all models
```{r, collapse=T, comment=NA, warning=F}
plot(Ridge_Chart, col = "blue")
plot(Lasso_Chart, add = TRUE, col = "red")
plot(GBM_Chart, add = TRUE, col = "black")
plot(XGBOOST_Chart, add = TRUE, col = "green")
legend("bottomright", legend=c("Ridge", "Lasso ","GBM","XGBOOST"),
       col=c( "blue","red","black","green","orange"), lty=1, 
       title="Line types", text.font=4, bg='lightblue')
```
